---
title: "pipeline"
output: html_document
---


```{r setup, include=FALSE}
library(tidyverse)
library(magrittr)
```

0. Install NetworKIN and in vitro database for IV-KEA
```{r}
source('R/install_networkin.R')
networkin_setup()
# ivkea_setup() #optional, should not have to be run (file given in github already with correct IDs)
```

1. Generate NetworKIN input from phosphoproteomic data
```{r}
# Human phosphoproteomic data
source('R/networkin_input.R')
networkin_input(
  phospho_file = "data/imported/phospho_human_LAM.xlsx",
  phosphosites_column = "PhosphoSite",
  log2_column = "LAM-Log2(Relapse/diagnostic",
  fdr_column = "pvalue LAM",
  experiment = "human",
  species = 'hsa'
)

# Mouse phosphoproteomic data
source('R/networkin_input.R')
networkin_input(
  phospho_file = "data/imported/phospho_mouse.xlsx",
  phosphosites_column = "ProteinID-Phospho:Site",
  log2_column = "Log2",
  fdr_column = "Adj. p-value",
  experiment = "mouse",
  species = 'mmu'
)
```

2. Run NetworKIN
__may not work on Windows__
```{r}
source('R/run_networkin.R')
run_networkin(
  input_file = "data/outputs/networKIN_input_human.res",
  blastall_location <- "~/blast-2.2.17/bin/blastall",
  experiment = "human"
)
```

3. NetworKIN predictions filtering
Predictions that have a score in the top 20% of all predictions are kept. Predictions are kept for all sites (significantly deregulated or not deregulated).
```{r}
source('R/filter_predictions.R')
filter_predictions(
  predictions_file = 'data/outputs/networKIN_output_human.tsv',
  experiment = 'human'
)
```

4. NetworKIN Fisher exact test
The following Fisher exact test can be performed with NetworKIN and IV-KEA:
- Up reg vs down reg
- Up reg vs everything (signif and not signif)
- Down reg vs everything (signif and not signif)
```{r}
source('R/select_top_predictions.R')
select_top_predictions(predictions_file = 'data/analyses/filtered_networkin_predictions_human.csv',
                phospho_cleaned_file = 'data/outputs/phospho_clean_human.csv',
                experiment = 'human')
source('R/perform_fisher_exact_test.R')
perform_Fisher_exact_test(top_predictions_file = 'data/analyses/top_predictions_human.csv',
                              experiment = 'networkin_human')
```

5. Generate volcano plots
Each Fisher exact test can be used to generate a volcano plot.
```{r}
source("R/make_volcano_plot.R")
make_volcano_plot(kinase_enrichment_file = 'data/analyses/kinase_enrichment_networkin_human.csv', 
                  odds_ratio = up_vs_down_odds_ratio,
                  FDR_cutoff = 0.05, 
                  FDR = up_vs_down_FDR,
                  graph_title = "Kinases prediction enrichment (NetworKIN) up vs down",
                  x_axis_title = "log2(odds ratio)",
                  save_path = "figures/up_down_volcano_plot_human.pdf")
make_volcano_plot(kinase_enrichment_file = 'data/analyses/kinase_enrichment_networkin_human.csv', 
                  odds_ratio = up_vs_tot_odds_ratio,
                  FDR_cutoff = 0.05, 
                  FDR = up_vs_down_FDR,
                  graph_title = "Kinases prediction enrichment (NetworKIN) up vs tot",
                  x_axis_title = "log2(odds ratio)",
                  save_path = "figures/up_tot_volcano_plot_human.pdf")
make_volcano_plot(kinase_enrichment_file = 'data/analyses/kinase_enrichment_networkin_human.csv', 
                  odds_ratio = down_vs_tot_odds_ratio,
                  FDR_cutoff = 0.05, 
                  FDR = up_vs_down_FDR,
                  graph_title = "Kinases prediction enrichment (NetworKIN) down vs tot",
                  x_axis_title = "log2(odds ratio)",
                  save_path = "figures/down_tot_volcano_plot_human.pdf")
```

6. Perform IV-KEA
```{r}
source('R/perform_ivkea.R')
perform_ivkea <- function(clean_phospho_file = 'data/outputs/phospho_clean_human.csv',
                          invitrodb_file = 'data/outputs/invitrodb.csv',
                          experiment = 'human'
                          )
```

Run Fisher Exact test
```{r}
source('R/perform_Fisher_exact_test.R')
perform_Fisher_exact_test(top_predictions_file = 'data/outputs/ivkea_predictions_human.csv',
                              experiment = 'ivkea_human')
```

Plot results on volcano plots
```{r}
source("R/make_volcano_plot.R")
make_volcano_plot(kinase_enrichment_file = 'data/analyses/kinase_enrichment_ivkea_human.csv', 
                  odds_ratio = up_vs_down_odds_ratio,
                  FDR_cutoff = 0.05, 
                  FDR = up_vs_down_FDR,
                  graph_title = "Kinases prediction enrichment (IV-KEA) up vs down",
                  x_axis_title = "log2(odds ratio)",
                  save_path = "figures/up_down_volcano_plot_human_ivkea.pdf")
make_volcano_plot(kinase_enrichment_file = 'data/analyses/kinase_enrichment_ivkea_human.csv', 
                  odds_ratio = up_vs_tot_odds_ratio,
                  FDR_cutoff = 0.05, 
                  FDR = up_vs_down_FDR,
                  graph_title = "Kinases prediction enrichment (IV-KEA) up vs tot",
                  x_axis_title = "log2(odds ratio)",
                  save_path = "figures/up_tot_volcano_plot_ivkea_human.pdf")
make_volcano_plot(kinase_enrichment_file = 'data/analyses/kinase_enrichment_ivkea_human.csv', 
                  odds_ratio = down_vs_tot_odds_ratio,
                  FDR_cutoff = 0.05, 
                  FDR = up_vs_down_FDR,
                  graph_title = "Kinases prediction enrichment (IV-KEA) down vs tot",
                  x_axis_title = "log2(odds ratio)",
                  save_path = "figures/down_tot_volcano_plot_ivkea_human.pdf")
```

