---
title: "How to use phosphogo"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{How to use phosphogo}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

This vignette highlights essential functions of `phosphogo` and their use.
For details on `phosphogo` installation, refer to the [GitHub repository](https://github.com/gaelfortin/phosphogo)
of `phosphogo`.

You can also use the web interface of pĥosphogo with:
```{r eval=FALSE}
library(phosphogo)
phosphogoApp()
```


First, load phosphogo and create a new folder to output all files:
```{r eval=FALSE}
library(phosphogo)
dir.create('myexperiment', showWarnings = FALSE) 
```


1. Generate NetworKIN and IV-KEA input from phosphoproteomic data
```{r eval=FALSE}
# Human phosphoproteomic data
phospho_input(
  phospho_file = "phospho_human.xlsx",
  phosphosites_column = "PhosphoSite",
  log2_column = "Log2",
  fdr_column = "Adj. Pvalue",
  species = 'hsa',
  output_folder = 'myexperiment/'
)

# Mouse phosphoproteomic data
phospho_input(
  phospho_file = "phospho_mouse.xlsx",
  phosphosites_column = "ProteinID-Phospho:Site",
  log2_column = "Log2",
  fdr_column = "Adj. Pvalue",
  species = 'mmu',
  output_folder = 'myexperiment/'
)
```

2. Run NetworKIN
NetworKIN predictions were pre-computed on all human phosphosites. Thus, phosphogo
fetches prediction for the given list of phosphosites.

```{r eval=FALSE}
run_networkin(
  input_file = "networKIN_input.res",
  output_folder = "myexperiment/"
)
```

3. Verify NetworKIN prediction scores distribution
```{r eval=FALSE}
networkin_qc(predictions_file = 'networkin_output.csv',
             output_folder = 'myexperiment/')
```

4. NetworKIN Fisher exact test
The following Fisher exact test can be performed with NetworKIN and IV-KEA:
- Up reg vs down reg
- Up reg vs everything (signif and not signif)
- Down reg vs everything (signif and not signif)
```{r eval=FALSE}
perform_Fisher_exact_test(predictions_file = 'networkin_output.csv',
                              predictions = "networkin",
                              output_folder = "myexperiment/",
                              FC_threshold = 1.2) 
```

5. Generate volcano plots
Each Fisher exact test can be used to generate a volcano plot.
```{r eval=FALSE}
make_volcano_plot(kinase_enrichment_file = 'kinase_enrichment_networkin.csv', 
                  odds_ratio = up_vs_down_odds_ratio,
                  FDR_cutoff = 0.05, 
                  FDR = up_vs_down_FDR,
                  graph_title = "Kinases prediction enrichment (NetworKIN) up vs down",
                  x_axis_title = "log2(odds ratio)",
                  output_folder = "myexperiment/",
                  file_name = "up_down_volcano_plot_networkin.pdf")
make_volcano_plot(kinase_enrichment_file = 'kinase_enrichment_networkin.csv', 
                  odds_ratio = up_vs_tot_odds_ratio,
                  FDR_cutoff = 0.05, 
                  FDR = up_vs_tot_FDR,
                  graph_title = "Kinases prediction enrichment (NetworKIN) up vs tot",
                  x_axis_title = "log2(odds ratio)",
                  output_folder = 'myexperiment/',
                  file_name = "up_tot_volcano_plot_networkin.pdf")
make_volcano_plot(kinase_enrichment_file = 'kinase_enrichment_networkin.csv', 
                  odds_ratio = down_vs_tot_odds_ratio,
                  FDR_cutoff = 0.05, 
                  FDR = down_vs_tot_FDR,
                  graph_title = "Kinases prediction enrichment (NetworKIN) down vs tot",
                  x_axis_title = "log2(odds ratio)",
                  output_folder = 'myexperiment/',
                  file_name = "down_tot_volcano_plot_networkin.pdf")
```

6. Perform IV-KEA
```{r eval=FALSE}
perform_ivkea(clean_phospho_file = 'phospho_clean.csv',
                          output_folder = 'myexperiment/')
```

Run Fisher Exact test
```{r eval=FALSE}
perform_Fisher_exact_test(predictions_file = 'ivkea_predictions.csv',
                              predictions = "ivkea",
                              output_folder = 'myexperiment/',
                              FC_threshold = 1.2)
```

Plot results on volcano plots
```{r eval=FALSE}
make_volcano_plot(kinase_enrichment_file = 'kinase_enrichment_ivkea.csv', 
                  odds_ratio = up_vs_down_odds_ratio,
                  FDR_cutoff = 0.05, 
                  FDR = up_vs_down_FDR,
                  graph_title = "Kinases prediction enrichment (IV-KEA) up vs down",
                  x_axis_title = "log2(odds ratio)",
                  output_folder = 'myexperiment/',
                  file_name = "up_down_volcano_plot_ivkea.pdf")
make_volcano_plot(kinase_enrichment_file = 'kinase_enrichment_ivkea.csv', 
                  odds_ratio = up_vs_tot_odds_ratio,
                  FDR_cutoff = 0.05, 
                  FDR = up_vs_tot_FDR,
                  graph_title = "Kinases prediction enrichment (IV-KEA) up vs tot",
                  x_axis_title = "log2(odds ratio)",
                  output_folder = 'myexperiment/',
                  file_name = "up_tot_volcano_plot_ivkea.pdf")
make_volcano_plot(kinase_enrichment_file = 'kinase_enrichment_ivkea.csv', 
                  odds_ratio = down_vs_tot_odds_ratio,
                  FDR_cutoff = 0.05, 
                  FDR = down_vs_tot_FDR,
                  graph_title = "Kinases prediction enrichment (IV-KEA) down vs tot",
                  x_axis_title = "log2(odds ratio)",
                  output_folder = 'myexperiment/',
                  file_name = "down_tot_volcano_plot_ivkea.pdf")
```
  
7. Compare enrichment tests from NetworKIN and IV-KEA
```{r eval=FALSE}
predictions_comparison(
  ivkea_enrichment_file = 'myexperiment/kinase_enrichment_ivkea.csv',
  networkin_enrichment_file = 'myexperiment/kinase_enrichment_networkin.csv',
  FDR_cutoff = 0.05, 
  graph_title = "Kinases prediction enrichment up vs down networkin vs iv-kea",
  output_folder = 'myexperiment/',
  file_name = "comp_plot.pdf"
)
```


8. Investigate deregulated kinases functions in cancer with DepMap


[DepMap](https://depmap.org/portal/) is a set of databases that aggregates dependency data for many cancer cell lineages. Phosphogo takes advantage of DepMap to identify deregulated kinases and there substrates that are important for the survival of cancers.
```{r}
# Import kinase predictions from NetworKIN and IVKEA
deregulated_kinases <- NULL
kinases_enrichment_files <- c('myexperiment/kinase_enrichment_networkin.csv', 'myexperiment/kinase_enrichment_ivkea.csv')
for(i in kinases_enrichment_files){ 
  deregulated_kinases <- c(deregulated_kinases,
                            read_csv(i) %>%
                               filter(up_vs_down_FDR < 0.05) %>%
                               pull(top_predicted_kinase))
                  }
deregulated_kinases <- unique(deregulated_kinases)
# Extract median dependency score of each kinase in cancer cell lineages
dependency_k_output <- dependency_kinases(kinases = deregulated_kinases, 
                                          dependency_threshold = -0.2,
                                          output_folder = 'myexperiment/')
# Extract substrates dependency scores of interesting kinases from dependency_kinases()
dependency_d_output <- dependency_data(dependency_kinases_output = "dependency_kinases.csv", 
                                       functional_threshold = 0.25,
                                       output_folder = 'myexperiment/')
# Generate the plot of dependency for interesting kinases from dependency_kinases()
dependency_g <- dependency_graph(dependency_data_output = "dependency_data.csv",
                                 output_folder = 'myexperiment/')
# Output dependency analysis results
dependency_k_output
dependency_g
dependency_d_output
```

