# 4 kinase enrichment approach

Up and downregulated proteome sets are compared to test any kinase enrichment. This test does not take the foldchange of each deregulated phosphosite into account.

# Extract top predictions for each phosphosite

```{r}
signif_predictions <- read_csv("data/analyses/filtered_predictions.csv", col_types = cols())
filtered_proteo <- read_csv("data/analyses/filtered_humanized_phospho_results.csv", col_types = cols())
```

## Select top predictions
For each phosphosite, keep only the prediction(s) with the top score

```{r}
top_predictions <- signif_predictions %>% 
  group_by(protein_phosphosite) %>% 
  filter(near(networkin_score, max(networkin_score))) %>% #near() is safer than == (see ?near)
  select(protein_phosphosite, "top_predicted_kinase" = `kinase_phosphatase_phospho-binding`)
```


## Merge proteo results and predictions
```{r}
proteo_predictions <- left_join(filtered_proteo, top_predictions, by = "protein_phosphosite") %>% 
  select(-MOD_RSD_mouse, -SITE_GRP_ID, -PROTEIN_mouse, -ACC_ID_mouse, -`SITE_+/-7_AA_mouse`)

write_csv(proteo_predictions, "data/analyses/proteo_predictions.csv")
```


# Subset proteo data into up and down regulated sets
```{r}
upreg_proteo <- proteo_predictions %>% 
  filter(is.na(top_predicted_kinase)==FALSE) %>% 
  filter(Log2 > 0) %>% 
  count(top_predicted_kinase)

downreg_proteo <- proteo_predictions %>% 
  filter(is.na(top_predicted_kinase)==FALSE) %>% 
  filter(Log2 < 0) %>% 
  count(top_predicted_kinase)
```

# Contingency table 
```{r}
contingency_table <- full_join(upreg_proteo, downreg_proteo, by = "top_predicted_kinase", suffix = c("_upreg", "_downreg"))

contingency_table %<>%  replace_na(replace = list(n_upreg = 0, n_downreg = 0))

contingency_table %<>% 
  remove_rownames() %>% 
  column_to_rownames("top_predicted_kinase") 
rm(upreg_proteo, downreg_proteo, proteo_predictions)
```


# Enrichment upregulated vs downregulated
## Fisher exact test
is kinase_1 enriched in the upregulated phosphosites or the downregulated phosphosites?
```{r}
source("R/fisher_exact_test.R")

contingency_table %<>% 
  mutate(fisher_res = map2(.x = n_upreg, .y = n_downreg, .f = fisher_exact_test, sum(contingency_table$n_upreg), sum(contingency_table$n_downreg))) %>% 
  separate(col = fisher_res, into = c("up_vs_down_pvalue", "up_vs_down_odds_ratio"), sep = ",", convert = TRUE)

```

## Adjust p-values (FDR)
```{r}
contingency_table %<>% bind_cols(up_vs_down_FDR =  p.adjust(contingency_table$up_vs_down_pvalue, method = "fdr")) 
```

# Enrichment up vs total and dow vs total
## Fisher exact test
```{r}
contingency_table %<>%
  mutate(n_tot = n_upreg + n_downreg) 
contingency_table %<>% 
  mutate(fisher_res_up = map2(.x = n_upreg, .y = n_tot, .f = fisher_exact_test, sum(contingency_table$n_upreg), sum(contingency_table$n_tot)),
         fisher_res_down = map2(.x = n_downreg, .y = n_tot, .f = fisher_exact_test, sum(contingency_table$n_downreg), sum(contingency_table$n_tot))) %>% 
  separate(col = fisher_res_up, into = c("up_vs_tot_pvalue", "up_vs_tot_odds_ratio"), sep=",", convert = TRUE) %>% 
  separate(col = fisher_res_down, into = c("down_vs_tot_pvalue", "down_vs_tot_odds_ratio"), sep=",", convert = TRUE)
```

## Adjust p-values (FDR)
```{r}
contingency_table %<>% bind_cols(
  up_vs_tot_FDR =  p.adjust(contingency_table$up_vs_tot_pvalue, method = "fdr"),
  down_vs_tot_FDR =  p.adjust(contingency_table$down_vs_tot_pvalue, method = "fdr"))%>% 
  rownames_to_column(var = "top_predicted_kinase")

write_csv(contingency_table, paste0("data/outputs/kinase_enrichment_LAM.csv"))
rm(fisher_exact_test)
```

# Plot kinase enrichment
## Volcano plot for up vs down
```{r}
source("R/make_volcano_plot.R")
vp_networkin <- make_volcano_plot(dataframe = contingency_table, 
                  odds_ratio = up_vs_down_odds_ratio,
                  FDR_cutoff = 0.05, 
                  FDR = up_vs_down_FDR,
                  graph_title = "Kinases prediction enrichment (NetworKIN) ERR",
                  x_axis_title = "log2(odds ratio)",
                  save_path = "figures/up_down_volcano_plot_ERR.pdf")
rm(list = ls())
```


